---
title: "TCGA.DATA R Package"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, eval=FALSE}
rmarkdown::render('README.Rmd', output_dir = '..')
```


This R Package allows to retrieve Gene Expression, Mutation and clinical data from [TCGA database](http://gdc-portal.nci.nih.gov/) (The Cancer Genome Atlas). It retrieves a single type of cancer at a time.

We publish diferent package in the [releases page](https://github.com/averissimo/tcga.data/releases) that allow to quickly use the datasets.

The genome expression datasets are already in a matrix format ready to be used. The data is in FPKM (Fragments Per Kilobase Million) format. Any additional normalization to use in models must be performed

## Package information

### How to use the dataset

1. Install `brca.data` by using `devtools` package. (`brca.data`, `prad.data` or `skcm.data`)

1. Load the library

1. Load the required datasets (one or more of the following)
    - `clinical`
    - `fpkm.per.tissue`
    - `fpkm.per.tissue.barcode`
    - `mutation`
    - `gdc`

#### Example

```{r install_brca.data, eval=FALSE}
# install the devtooks library
install.packages('devtools')
# The library can also be loaded and use the function install_git without 'devtools::' prefix
devtools::install_url('https://github.com/averissimo/tcga.data/releases/download/2016.12.15-brca/brca.data_1.0.tar.gz')
#
# Load the brca.data package
library(brca.data)
# start using the data, for example the tissue data
data(fpkm.per.tissue)
# tissue is now in the enviromnet and will be loaded on the first
#  time it is used. For example:
names(fpkm.per.tissue)
```

## Ackowledgements

This package was developed primarily by *[André Veríssimo](http://web.tecnico.ulisboa.pt/andre.verissimo/)* with support from *Marta Lopes* and *[Susana Vinga](http://web.tecnico.ulisboa.pt/susanavinga/)*

This work was supported by:

- [FCT](www.fct.pt), through IDMEC, under LAETA, projects *(UID/EMS/50022/2013)*;
  - Susana Vinga acknowledges support by program
 Investigador FCT *(IF/00653/2012)* from [FCT](www.fct.pt), co-funded by the European Social Fund *(ESF)* through the Operational Program Human Potential *(POPH)*;
  - André Veríssimo acknowledges support from [FCT](www.fct.pt) *(SFRH/BD/97415/2013)*.
